---
title: "Figure 1"
author: "Kenny Huang & Reha Mathur"
output: html_notebook
---

## installing packages and loading libraries 

```{r}
library("devtools")
install_github("jw156605/SLICER")
```

```{r}
library(Seurat)
library(readr)
library(tidyverse)
library(HGC) 
library(latticeExtra)
library(lattice) 
library(RColorBrewer)
library(clusterProfiler)
library(org.Mm.eg.db)
library(AnnotationDbi)
library(ComplexHeatmap)
library(dendextend)
library('fastcluster')

# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("clusterProfiler")
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("AnnotationDbi")
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("org.Mm.eg.db")
```

## reading in the expression data from the paper 

```{r}
## Kenny wrote this 
exp = read_csv("normalized_data.csv")
exp = discard(exp,~all(is.na(.) |.  == ""))
exp = distinct(exp, `...1`, .keep_all = TRUE)
# filtering out the reprogramming transcription factors 
exp = filter(exp, !`...1`%in% c("Mef2c", "Gata4", "Tbx5"))
exp = column_to_rownames(exp, var="...1")
exp_log2 = log2(exp)
exp_log2[exp_log2 < 0] <- 0
exp_log2_mat = as.matrix(exp_log2)
```

## creating seurat object 

```{r}
# Kenny wrote this 
s = CreateSeuratObject(counts = exp)
s = SetAssayData(object = s, slot = "scale.data", new.data = exp_log2_mat)
s <- FindVariableFeatures(s, selection.method = "vst", nfeatures = 10000)
```

## running PCA on seurat object 

```{r}
# Kenny wrote this 
s = RunPCA(s, features  = VariableFeatures(s))
```

## extracting top 400 by max weights

```{r}
# Kenny wrote this 
our_genes = s[["pca"]]@feature.loadings %>% 
  data.frame() %>% 
  dplyr::select(PC_1, PC_2, PC_3, PC_4) %>% 
  mutate(PC_1 = abs(PC_1), PC_2 = abs(PC_2), PC_3 = abs(PC_3), max = pmax(PC_1, PC_2,PC_3), sum = PC_1 + PC_2 + PC_3) %>% 
  slice_max(order_by = max, n = 400) %>% 
  row.names() %>% 
  sort()
capture.output(our_genes, file = "top_genes.txt", row.names = FALSE)
```

## reading in the genes that they said were in the top 400 for comparison 
```{r}
# Kenny wrote this 
correct_genes = read_csv("correct_genes.csv", col_names = FALSE)
(our_genes %in% correct_genes$X1) %>% mean()
```

```{r}
# Kenny wrote this 
top_genes = exp_log2[our_genes,] %>% data.frame()
top_genes = t(top_genes) 
top_genes_scale = scale(top_genes)
```

# PCA Graphs

```{r}
# Kenny wrote this 
states = readr::read_csv("paper_cell_states.csv")
states = column_to_rownames(states, var="Cell_ID")
states = states[rownames(top_genes_scale),]
stage = states$stage
stage[stage == "IFib"] = "Fib"
stage[stage == "AFib"] = "Fib"
stage[stage == "IiCM"] = "iCM"
stage[stage == "AiCM"] = "iCM"
```

```{r}
# Kenny wrote this 
states2 = readr::read_csv("categorized_cells_1.csv")
states2 = column_to_rownames(states2, var="name")
stage2 = states2[rownames(top_genes_scale),]
stage2 = substr(stage2, 4, 8)
```

```{r}
# Kenny wrote this 
genes_pca <- prcomp(t(top_genes_scale), center = TRUE, scale. = TRUE)
genes_pca %>% summary()
ggbiplot::ggbiplot(genes_pca, var.axes = FALSE, obs.scale = 1, var.scale = 1, group=gene_labeling$label,
     varname.size = 3, labels.size=3) +
scale_color_manual(name="Gene Type", values=c("blue", "orange", "red"))
```

```{r}
# Kenny wrote this 
cells_pca <- prcomp(top_genes_scale, center = TRUE, scale. = TRUE)
cells_pca %>% summary()
ggbiplot::ggbiplot(cells_pca, var.axes = FALSE, obs.scale = 1, var.scale = 1, group=stage2,
     varname.size = 3, labels.size=3) +
scale_color_manual(name="Cell Type", values=c("blue", "red", "purple", "lightgreen"))
```

# fixed clustering
```{r}
# Kenny wrote this 
row_dend = hclust(dist(top_genes_scale)) # row clustering
col_dend = hclust(dist(t(top_genes_scale))) # column clustering
```

# Gene ontology analysis 

```{r}
# Reha wrote this 
clusters <- cutree(col_dend, k = 3)
cluster1 = colnames(top_genes_scale)[clusters == 1]
GO_results_cluster1 <- enrichGO(gene = cluster1, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP")
as.data.frame(GO_results_cluster1)

cluster2 = colnames(top_genes_scale)[clusters == 2]
GO_results_cluster2 <- enrichGO(gene = cluster2, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP")
as.data.frame(GO_results_cluster2)

cluster3 = colnames(top_genes_scale)[clusters == 3]
GO_results_cluster3 <- enrichGO(gene = cluster3, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP")
as.data.frame(GO_results_cluster3)

```

```{r}
# Reha wrote this 
dd.col = as.dendrogram(col_dend)
col.ord = order.dendrogram(dd.col)
fit <- plot(barplot(GO_results_cluster1, showCategory = 6))
fit <- plot(barplot(GO_results_cluster2, showCategory = 6))
fit <- plot(barplot(GO_results_cluster3, showCategory = 6))

gene_labeling = data.frame(matrix(ncol = 1, nrow = length(colnames(top_genes_scale))))
colnames(gene_labeling) = "label"
rownames(gene_labeling) = colnames(top_genes_scale)
gene_labeling[cluster1,] = "fibroblast"
gene_labeling[cluster3,] = "cardiomyocyte"
gene_labeling[cluster2,] = "cell cycle"

gene_labeling[col.ord, ]

```

```{r}
# Reha wrote this 
library("stringr") 

cluster1_df = as.data.frame(GO_results_cluster1)
cluster1_df = cluster1_df[order(cluster1_df[,7]), ]
p<-ggplot(data=cluster1_df[1:6,], aes(x=Description, y=Count/152*100, fill=qvalue)) +
  geom_bar(stat="identity") + 
  coord_flip()+scale_fill_gradient(high = "#2c2f92", low = "#a0a2e2")+ theme(text = element_text(size = 50)) +      theme(legend.spacing.y = unit(1.0, 'cm'))  +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30)) + theme(legend.text = element_text(size = 30))  + labs( y = "Genes Involved (%)") + theme(legend.key.size = unit(2, 'cm'))
p

cluster2_df = as.data.frame(GO_results_cluster2)
cluster2_df = cluster2_df[order(cluster2_df[,7]), ]
p<-ggplot(data=cluster2_df[1:6,], aes(x=Description, y=Count/170 *100, fill=qvalue)) +
  geom_bar(stat="identity") + 
  coord_flip()+scale_fill_gradient(high = "#ffcc00", low = "#fff0b3")+ theme(text = element_text(size = 50)) +theme(legend.spacing.y = unit(1.0, 'cm'))  +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30)) + theme(legend.text = element_text(size = 30))  + labs( y = "Genes Involved (%)") + theme(legend.key.size = unit(2, 'cm'))
p
cluster3_df = as.data.frame(GO_results_cluster3)
cluster3_df = cluster3_df[order(cluster3_df[,7]), ]
p<-ggplot(data=cluster3_df[1:6,], aes(x=Description, y=Count/78 * 100, fill=qvalue)) +
  geom_bar(stat="identity") + 
  coord_flip()+scale_fill_gradient(high = "#ff3300", low = "#ffad99")+ theme(text = element_text(size = 50))   +theme(legend.spacing.y = unit(1.0, 'cm'))  +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30)) + theme(legend.text = element_text(size = 30))  + labs( y = "Genes Involved (%)") + theme(legend.key.size = unit(2, 'cm'))
p
```


# heatmap of results 

```{r}
# Kenny wrote this 
Heatmap(t(top_genes_scale), name = "test", 
        cluster_rows = color_branches(col_dend, k = 3),
        cluster_columns = color_branches(row_dend, k = 30, col = c("darkred", "forestgreen", "orange", "blue")), show_row_names = FALSE, show_column_names = FALSE)
```
# active cells categorization for other analysis
```{r}
# Kenny wrote this 
cell_clusters <- cutree(row_dend, k = 2)
active_cells = rownames(top_genes_scale)[cell_clusters == 1]
```


```{r}
# Kenny wrote this 
cell_categorizations = cutree(row_dend, k = 30, order_clusters_as_data = FALSE)
```

# categorizing all of the cells for further analysis
```{r}
# Kenny wrote this 
CCA_iCM = names(cell_categorizations[cell_categorizations == 1 | cell_categorizations == 2])
CCA_Fib = names(cell_categorizations[cell_categorizations == 3 | cell_categorizations == 4 | cell_categorizations == 5 | cell_categorizations == 6 | cell_categorizations == 7 | cell_categorizations == 8 | cell_categorizations == 9 ])
CCA_iFib = names(cell_categorizations[cell_categorizations == 10 | cell_categorizations == 11 | cell_categorizations == 12 | cell_categorizations == 13 | cell_categorizations == 14 ])
CCI_piCM = names(cell_categorizations[cell_categorizations == 15  | cell_categorizations == 16 | cell_categorizations == 17 | cell_categorizations == 18 | cell_categorizations == 23])
CCI_iCM = names(cell_categorizations[cell_categorizations == 19  | cell_categorizations == 20])
CCI_iFib = names(cell_categorizations[cell_categorizations == 21  | cell_categorizations == 22 |  cell_categorizations == 26 | cell_categorizations == 27 |  cell_categorizations == 28|  cell_categorizations == 29])
CCI_Fib = names(cell_categorizations[cell_categorizations == 24  | cell_categorizations == 25| cell_categorizations == 30 ])
```

```{r}
# Kenny wrote this 
cell_categorized = data.frame(name= rownames(top_genes_scale), cat = "CCI", stage = "FIB")
cell_categorized[cell_categorized$name %in% CCA_Fib,2] = "CCAFib"
cell_categorized[cell_categorized$name %in% CCA_iFib,2] = "CCAiFib"
cell_categorized[cell_categorized$name %in% CCA_iCM,2] = "CCAiCM"
cell_categorized[cell_categorized$name %in% CCI_iFib,2] = "CCIiFib"
cell_categorized[cell_categorized$name %in% CCI_Fib,2] = "CCIFib"
cell_categorized[cell_categorized$name %in% CCI_piCM,2] = "CCIipCM"
cell_categorized[cell_categorized$name %in% CCI_iCM,2] = "CCIiCM"

cell_categorized[cell_categorized$name %in% CCA_Fib,3] = "Fib"
cell_categorized[cell_categorized$name %in% CCA_iFib,3] = "iFib"
cell_categorized[cell_categorized$name %in% CCA_iCM,3] = "iCM"
cell_categorized[cell_categorized$name %in% CCI_iFib,3] = "iFib"
cell_categorized[cell_categorized$name %in% CCI_Fib,3] = "Fib"
cell_categorized[cell_categorized$name %in% CCI_piCM,3] = "piCM"
cell_categorized[cell_categorized$name %in% CCI_iCM,3] = "iCM"
```

```{r}
write.csv(cell_categorized, "categorized_cells.csv" , row.names = FALSE)
```







